---
output:
pdf_document: default
html_document: default
---
Explore White Wine Quality 
========================================================

### By Ruchita Maheshwary

In this report, I analyzed the data set for around 5000 white wines.

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}

#install.packages("wesanderson")
#install.packages("corrplot")
#install.packages("RColorBrewer")
#install.packages("ppcor")
library(ggplot2)
library(data.table)
library(dplyr)
library(gridExtra)
library(GGally)
library(scales)
library(memisc)
library(wesanderson)
library(corrplot)
library(RColorBrewer)
library(ppcor)
library("ggthemes")
library("scales")


```

```{r global_options, include=FALSE}

knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)

```


```{r style}
#Styling the plots
theme_update(axis.title=element_text(size=12),axis.text=element_text(size=10),
             plot.title=element_text(size=14))

```


```{r Load_the_Data}
# Load the Data
white_wine_df <-read.csv('wineQualityWhites.csv')
white_wine_df$X <- NULL

```

This data set is used to model wine preferences by data mining from 
physicochemical properties. In the dataset, the inputs include objective tests 
(e.g. PH values) and the output is based on sensory data (median of at least 3 
evaluations made by wine experts). Each expert graded the wine quality between 0
(very bad) and 10 (very excellent). 

**Brief description about the dataset variables:**

* fixed acidity: most acids involved with wine or fixed or nonvolatile (do not 
evaporate readily)
* volatile acidity: the amount of acetic acid in wine, which at too high of 
levels can lead to an unpleasant, vinegar taste
* citric acid: found in small quantities, citric acid can add 'freshness' and 
flavor to wines
* residual sugar: the amount of sugar remaining after fermentation stops, it's 
rare to find wines with less than 1 gram/liter and wines with greater than 45 
grams/liter are considered sweet
* chlorides: the amount of salt in the wine
* free sulfur dioxide: the free form of SO2 exists in equilibrium between 
molecular SO2 (as a dissolved gas) and bisulfite ion; it prevents microbial 
growth and the oxidation of wine
* total sulfur dioxide: amount of free and bound forms of S02; in low 
concentrations, SO2 is mostly undetectable in wine, but at free SO2 
concentrations over 50 ppm, SO2 becomes evident in the nose and taste of wine
* density: the density of water is close to that of water depending on the 
percent alcohol and sugar content
* pH: describes how acidic or basic a wine is on a scale from 0 (very acidic) 
to 14 (very basic); most wines are between 3-4 on the pH scale
* sulphates: a wine additive which can contribute to sulfur dioxide gas (S02) 
levels, wich acts as an antimicrobial and antioxidant
* alcohol: the percent alcohol content of the wine
* quality (score between 0 and 10): Output variable (based on sensory data)

# Univariate Plots Section

Below is a brief summary of the data:

```{r}
dim(white_wine_df)

str(white_wine_df)

summary(white_wine_df)

```

Here in this section I would like to explore the significance of couple of
variables in the data set.


```{r univariate_plot_fixed_acidity}

summary(white_wine_df$fixed.acidity)

# Plot for fixed acidity

p1<-qplot(x=fixed.acidity,data=white_wine_df,binwidth=0.1,
          xlab = "Fixed Acidity")+
  scale_x_continuous(limits=c(3,10.5),breaks=seq(3,10.5,3))+
  ggtitle("Fixed Acidity per g / dm^3")

p2<-qplot(y=fixed.acidity,x=1,data=white_wine_df,
          geom = 'boxplot',xlab = "Count",
          ylab = "Fixed acidity")+
  ggtitle("Fixed Acidity per g / dm^3")

grid.arrange(p1,p2,ncol=2)

```

White wine's fixed acidity plot is a normal distribution with highest 
concentration of fixed acidity at 6.8. Around 50% of wines have fixed acidity 
more than 6.8 and around top 25% lie in the range 6.8-7.3 with an outlier at
14.2.


```{r univariate_plot_volatile_acidity}

summary(white_wine_df$volatile.acidity)

p1<-qplot(x=volatile.acidity,data=white_wine_df,
      binwidth=0.01,xlab="Volatile Acidity")+
  scale_x_continuous(limits=c(0,0.75),breaks = seq(0,0.75,0.5))+
  ggtitle("Volatile Acidity per g / dm^3")

p2<-qplot(y=volatile.acidity,x=1,data=white_wine_df,
          geom = 'boxplot',xlab = "Count",
          ylab = "Volatile acidity")+
  ggtitle("Volatile Acidity per g / dm^3")

grid.arrange(p1,p2,ncol=2)

```

The summary statistics show that maximum volatile acidity is at 0.26. 
Around 50% of wines have volatile acidity more than 0.26 and around 
top 25% lie in the range 0.26-0.32 with an outlier at
1.1.



```{r univariate_plot_citric_acid}

summary(white_wine_df$citric.acid)

p1<-qplot(x=citric.acid,data=white_wine_df,binwidth=0.01,xlab = "Citric Acid")+
  scale_x_continuous(limits=c(0,0.8),breaks=seq(0,0.8,0.2))+
  ggtitle("Citric acid concentration per g / dm^3")

p2<-qplot(y=citric.acid,x=1,data=white_wine_df,geom = 'boxplot',xlab = "Count",
          ylab = "Citric Acid")+
  ggtitle("Citric acid concentration per g / dm^3")

grid.arrange(p1,p2,ncol=2)

```

The plot has long tailed which after re - adjusting the axis could be normally 
distributed. Here the median of citric acid concentration is 0.32 g/dm^3 with an
outlier at around 1.7.

```{r univariate_plot_residual_sugar}

summary(white_wine_df$residual.sugar)

p1<-qplot(x=residual.sugar,data=white_wine_df,
      binwidth=0.1,xlab = "Residual Sugar")+
  scale_x_continuous(limits=c(0,20),breaks = seq(0,20,5))+
  ggtitle("Residual Sugar Plot")

p2<-qplot(y=residual.sugar,x=1,data=white_wine_df,geom = 'boxplot',xlab = "Count",
          ylab = "Residual Sugar")+
  ggtitle("Residual Sugar Plot")

grid.arrange(p1,p2,ncol=2)

```

Here the median residual sugar is 5.2 g/dm^3 with an outlier at 65.8.
We can perform readjust the plot to get a better understanding of it.
This visualisation  of residual sugar appears skewed to the right with a 
high peak at around 3.


```{r Univariate_plot_chlorides}

summary(white_wine_df$chlorides)

p1<-ggplot(aes(x=chlorides),data=white_wine_df,binwidth=0.05)+
  geom_histogram()+
  scale_x_continuous(limits=c(0,0.1),breaks=seq(0,0.1,0.05))+
  ggtitle("Chloride content per g / dm^3")

p2<-qplot(y=chlorides,x=1,data=white_wine_df,geom = 'boxplot',xlab = "Count",
          ylab = "Chloride")+
  ggtitle("Chloride content per g / dm^3")

grid.arrange(p1,p2,ncol=2)


```

This visualization has a long tail with an outlier at around 0.34.This visualization 
appears slightly bimodal with rough peaks at around 0.03,0.05



```{r Univariate_free_sulfur_dioxidet}

summary(white_wine_df$free.sulfur.dioxide)

p1<-qplot(x=free.sulfur.dioxide,data=white_wine_df,binwidth=2,
      xlab = "Free sulfur dioxide ")+
  scale_x_continuous(limits = c(0,100),breaks = seq(0,100,10))+
  ggtitle("Free sulfur dioxide per g/dm^3")

p2<-qplot(y=free.sulfur.dioxide,x=1,data=white_wine_df,geom = 'boxplot',
          xlab = "Count",
          ylab = "Free sulfur dioxide")+
  ggtitle("Free sulfur dioxide per g/dm^3")

grid.arrange(p1,p2,ncol=2)

```


This graph has a median of free sulfur dioxide of 34 g/dm^3 and have an outlier
at 289. On re-adjusting the axis to obtain an understanding of the shape of 
the plot.After readjusting the axis to remove the outlier and after 
limit the range of x-axis to 0-100 g/dm^3, the plot is clearly a normal 
distribution. 

```{r Univariate_total_sulfur_dioxide}

summary(white_wine_df$total.sulfur.dioxide)

p1<-qplot(x=total.sulfur.dioxide,data=white_wine_df,binwidth=5,
      xlab="Total sulfur dioxide")+
  scale_x_continuous(limits = c(0,270),breaks = seq(0,270,50))+
  ggtitle("Total sulfur dioxide per g/dm^3")

p2<-qplot(y=total.sulfur.dioxide,x=1,data=white_wine_df,geom = 'boxplot',
          xlab = "Count",
          ylab = "Total sulfur dioxide")+
  ggtitle("Total sulfur dioxide per g/dm^3")

grid.arrange(p1,p2,ncol=2)

```

This plot is normally distributed with an outlier at 440 g/dm^3 with maximum
concentration of sulfur dioxide at around 134 g/dm^3.


```{r Univariate_plot_density}

summary(white_wine_df$density)

p1<-qplot(x=density,data=white_wine_df,binwidth=0.0005)+
  scale_x_continuous(limits = c(0.987,1.01),breaks = seq(0.987,1.01,0.01))+
  ggtitle("Density per g/cm^3")

p2<-qplot(y=density,x=1,data=white_wine_df,geom = 'boxplot',
          xlab = "Count",
          ylab = "Density")+
  ggtitle("Density per g/cm^3")

grid.arrange(p1,p2,ncol=2)

```

This plot had large number of outliers after 1.01 g/cm^3 with max being 1.0390
which had been removed after re-adjusting the x-axis. Now the plot appears 
to be normally distributed.The median of density had been at 0.9937 g/cm^3.   

```{r Univariate_plot_pH}

summary(white_wine_df$pH)

qplot(x=pH,data=white_wine_df,binwidth=0.02)+
  scale_x_continuous(limits = c(2.7,4),breaks = seq(2.7,4,0.5))+
  ggtitle("pH level")

```

This plot of pH level in the white wines is normally distributed with
the median at 3.188. Around 50% of white wines had pH level of more than 3.188
with the maximum being at 3.820.


```{r Univariate_plot_sulphates}

summary(white_wine_df$sulphates)

p1<-qplot(x=sulphates,data=white_wine_df,binwidth=0.02)+
  scale_x_continuous(limits = c(0.2,0.8),breaks = seq(0,0.8,0.25))+
  ggtitle("Sulphate per g/dm^3")

p2<-qplot(y=sulphates,x=1,data=white_wine_df,geom = 'boxplot',
          xlab = "Count")+
  ggtitle("Sulphate per g/dm^3")

grid.arrange(p1,p2,ncol=2)

```

This plot is skewed to the right with a long tail and having an outlier at 1.08.
The median of sulphate concentration is 0.47 g/dm^3.

```{r Univariate_plot_alcohol}

summary(white_wine_df$alcohol)

p1<-qplot(x=alcohol,data=white_wine_df,binwidth=0.1)+
  scale_x_continuous(limits = c(8,15),breaks = seq(8,15,1))+
  ggtitle("Alcohol content")

p2<-qplot(y=alcohol,x=1,data=white_wine_df,geom = 'boxplot',
          xlab = "Count")+
  ggtitle("Alcohol content")

grid.arrange(p1,p2,ncol=2)

```

This plot appears to be skewed towards the right having mean of 10.51 and
and an outlier at 14.20.


```{r Univariate_plot_quality}

summary(white_wine_df$quality)

ggplot(aes(x=quality),data=white_wine_df)+
  geom_bar()+
  scale_x_continuous(breaks=seq(3,9,1))+
  ggtitle("Plotting wine quality")

```


This bar plot shows that the median of wine quality has been 6 forming around
44.8% of the dataset with the best quality of wine having quality as 9 which 
forms around 1.02% of the dataset.



# Univariate Analysis

### What is the structure of your dataset?

There are 4898 observations of white wine in this dataset with 13 variables.
The first variable 'X' is a unique identifier for each record so it is ignored
in my analysis. Other variables like 	fixed.acidity, volatile.acidity,
citric.acid,residual.sugar,chlorides,free.sulfur.dioxide,total.sulfur.dioxide,
density,pH,sulphates,alcohol are all numeric variables with	quality 
being an integer variable.

### What is/are the main feature(s) of interest in your dataset?

My main point of interest is determing the factors which determine the quality
of white wine.

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

I would need to compute the correlation of each of the variables like 
fixed.acidity, volatile.acidity,citric.acid,residual.sugar,chlorides,
free.sulfur.dioxide,total.sulfur.dioxide, density,pH,sulphates,alcohol with the 
quality to determine which variables influence the quality and hence the 
taste of white wine. I will explore these relationships in the later sections
of my report.


### Did you create any new variables from existing variables in the dataset?

```{r Univariate_wine_quality_calculation }

#Adding extra variable for rating of the wine quality with 3-5 for bad,
# 6 for average and 7-9 as good

white_wine_df$rating <- NULL
white_wine_df$rating <-ifelse(white_wine_df$quality>=3 &
                                white_wine_df$quality<=5,
                              'Bad',
                              ifelse(white_wine_df$quality==6 ,
                                     'Average','Good'))

white_wine_df$rating <- factor(white_wine_df$rating)
white_wine_df$rating  <- factor(white_wine_df$rating , 
                                levels = c("Bad", "Average", "Good"))

#Calculate total acidity as sum of fixed acidity and volatile acidity
white_wine_df$total.acidity <- NULL
white_wine_df$total.acidity <- white_wine_df$fixed.acidity + 
                      white_wine_df$volatile.acidity

#Determine wine type on the basis of sweetness
white_wine_df$wine.type <-ifelse(white_wine_df$residual.sugar<10,'Dry',
                                 ifelse(white_wine_df$residual.sugar>30,
                                        'Sweet','Medium'))

#Group wines in buckets on the basis of citric acid concentration
white_wine_df$citric_acid_bucket <- cut(white_wine_df$citric.acid,
                                        c(0,0.3,0.6,1.7))

#Group wines in buckets on the basis of total sulfur dioxide concentration
white_wine_df$total_sulfur_dioxide_bucket <- 
  cut(white_wine_df$total.sulfur.dioxide, c(0,120,210,440))

#Group wines on the basis of alcohol content
white_wine_df$alcohol_content <-ifelse(white_wine_df$alcohol>=0 &
                                         white_wine_df$alcohol<12.5,'Low',
                                       ifelse(white_wine_df$alcohol>=12.5 &
                                                white_wine_df$alcohol<=13.5,
                                              'Moderately Low','High'))
white_wine_df$alcohol_content  <- factor(white_wine_df$alcohol_content , 
                                levels = c("Low", "Moderately Low", "High"))

#Chloride to sulfate ratio

white_wine_df$chloride_sulfate_ratio <- 
  ifelse((white_wine_df$chlorides/white_wine_df$sulphates)<0.50,
         'Very Bitter','Bitter')

```


I have added the following variable in the dataset :<br/>
* **rating** : which is calculated on the basis of the quality of the white 
wine. With 3-5 for "Bad",  6 for "Average" and 7-9 as "Good" rating. I have chosen
these scales to somewhat equally distribute the wine data set into various groups.
This factored variable will help me in later investigating the relationship of 
variables with the wine quality<br/>
* **total.acidity** : Calculated as sum of fixed and volatile acidity<br/>
* **wine.type** : Classifies wine on the basis of sugar content, with wines 
less than 10 sugar as "Dry", more than 30 as "sweet" and the others as "Medium"
* **citric_acid_bucket** : Groups citric acid into buckets of [0-0.3],(0.3-0.6],
(0.6,1.7]<br/>
* **total_sulfur_dioxide_bucket** :  Groups total sulfur dioxide into buckets of
[0-120],(120-210],(210,440]<br/>
* **alchol_content** :  Classifies wine on the basis of alcohol content, with 
wines content less than 12.5  as "Low", between 12.5 to 13.5  as "Moderately 
Low" and the others as "High"<br/>
* **chloride_sulfate_ratio** :  Classifies the proportion of chloride with 
sulfates with proportion less than 0.5 as "Very Bitter" and others as "Bitter"
<br/>

Since the wine rating variable will be used in my further analysis, I want
to plot a visualization depicting the distribution of wines in the different
"rated" buckets

```{r univariate_rating_distribution}

summary(white_wine_df$rating)

ggplot(aes(x=rating),data=white_wine_df)+
  geom_bar()+
  ggtitle("Plotting wine rating variable")

```

From this visualization it is clear that maximum wines have "average" rating(quality=6),
followed by "bad"(quality between 3 to 5) and then "good"(quality between 7 to 9)

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?

The fixed acidity,chloride and alcohol variables had to be log transformed
to make the plot more normally distributed. Other than that citric acid, 
sulfur dioxide and residual sugar plots x- axis had to  be re adjusted 
to have a better understanding of the visualization.


# Bivariate Plots Section


Here in this section I would like to explore the relationship among couple of
variable pairs which have a strong/weak relationship among them.


Analysis of relationship between the variables 

```{r  include=FALSE,correlation_values}


cor.test(white_wine_df$quality, white_wine_df$citric.acid,method = 'pearson')
#-0.009209091
cor.test(white_wine_df$quality, white_wine_df$free.sulfur.dioxide,
         method = 'pearson')
#0.008158067
cor.test(white_wine_df$quality, white_wine_df$fixed.acidity,method = 'pearson')
#-0.1136628
cor.test(white_wine_df$quality, white_wine_df$volatile.acidity,
         method = 'pearson')
#-0.194723
cor.test(white_wine_df$quality, white_wine_df$residual.sugar,method = 'pearson')
#-0.09757683
cor.test(white_wine_df$quality, white_wine_df$chlorides,method = 'pearson')
#-0.2099344
cor.test(white_wine_df$quality, white_wine_df$total.sulfur.dioxide,
         method = 'pearson')
#-0.1747372
cor.test(white_wine_df$quality, white_wine_df$density,method = 'pearson')
#-0.3071233 
cor.test(white_wine_df$quality, white_wine_df$pH,method = 'pearson')
#0.09942725
cor.test(white_wine_df$quality, white_wine_df$sulphates,method = 'pearson')
#0.05367788
cor.test(white_wine_df$quality, white_wine_df$alcohol,method = 'pearson')
#0.4355747

cor.test(white_wine_df$fixed.acidity, 
         white_wine_df$volatile.acidity,method = 'pearson')
#-0.02269729
cor.test(white_wine_df$alcohol, white_wine_df$citric.acid,
         method = 'pearson')
#-0.07572873
cor.test(white_wine_df$free.sulfur.dioxide, 
         white_wine_df$total.sulfur.dioxide,method = 'pearson')
#0.615501
cor.test(white_wine_df$citric.acid, white_wine_df$pH,method = 'pearson')
#-0.1637482
cor.test(white_wine_df$citric.acid, white_wine_df$residual.sugar,
         method = 'pearson')
#0.09421162 
cor.test(white_wine_df$alcohol, white_wine_df$residual.sugar,
         method = 'pearson')
#-0.4506312
cor.test(white_wine_df$fixed.acidity, white_wine_df$citric.acid,
         method = 'pearson')
#0.2891807
cor.test(white_wine_df$chlorides, white_wine_df$sulphates,
         method = 'pearson')
#0.01676288
cor.test(white_wine_df$chlorides, white_wine_df$residual.sugar,
         method = 'pearson')
#0.08868454 
cor.test(white_wine_df$chlorides, white_wine_df$density,method = 'pearson')
#0.2572113
cor.test(white_wine_df$alcohol, white_wine_df$density,method = 'pearson')
#-0.7801376

cor.test(white_wine_df$pH, white_wine_df$density,method = 'pearson')
#-0.09359149

cor.test(white_wine_df$pH, white_wine_df$fixed.acidity,
         method = 'pearson')
#-0.4258583

cor.test(white_wine_df$residual.sugar, white_wine_df$density,
         method = 'pearson')
#0.8389665

cor.test(white_wine_df$density, white_wine_df$total.sulfur.dioxide,
         method = 'pearson')
#0.5298813

cor.test(white_wine_df$alcohol,
         white_wine_df$total.sulfur.dioxide,method = 'pearson')
#-0.4488921 

cor.test(white_wine_df$chlorides, white_wine_df$alcohol,
         method = 'pearson')
#-0.3601887

```


```{r fig.height=15,fig.width=20,Bi_correlation_plot}

#Remove headers of variables in new data set which are not required for 
#correlation
correlation_headers <- names(white_wine_df) %in% c("rating","total.acidity",
                                                   "wine.type",
"citric_acid_bucket","total_sulfur_dioxide_bucket",
"alcohol_content","chloride_sulfate_ratio") 

wine_data <- white_wine_df[!correlation_headers]

 ggscatmat(wine_data,alpha=0.2) +
  theme_grey(base_size=7)+
   ggtitle('Correlation values for pairs of variables') +
   theme(plot.title = element_text(size = 16))



```


Numeric correlations among the variables

```{r fig.height=10,Bi_corr_plot_numeric}


corrplot(cor(wine_data), method="number")

```

From the above plot the top most positive and negative correlation 
coefficients, r based on strength of association[1] are 

* Residual sugar: density = 0.8389665
* Free sulfur dioxide: Total sulfur dioxide=0.615501
* Alcohol: quality = 0.4355747
* Total sulfur dioxide: density = 0.5298813
* Fixed acidity:Citric acid=0.2891807
* Chlorides:Density=0.2572113
* Density:quality= -0.3071233
* Alcohol:Residual sugar=-0.4506312
* Alcohol:Density= -0.7801376
* Fixed.acidity:pH= -0.4258583
* Alchohol: Total sulfur dioxide = -0.4488921 
* Alcohol: Chloride = -0.3601887

I will exploring some of these relationships in this section of the report.


```{r bivariate_sugar_density}

ggplot(aes(x=density,y=residual.sugar),data=white_wine_df)+
  geom_point(alpha=1/10,position=position_jitter(h=0))+
  xlim(0.985,1.005)+ylim(0,30)+
  ylab("Residual sugar")+
  xlab("Density")+
  geom_smooth(method='lm',color='red')+
  ggtitle("Residual sugar v/s density")
```

This plot shows that as residual sugar in white wine increases, the density
increases.


```{r bivariate_sulfur_dioxide}

ggplot(aes(x=free.sulfur.dioxide,y=total.sulfur.dioxide),data=white_wine_df)+
  geom_point(alpha=1/10,position=position_jitter(h=0.1))+
  xlim(0,100)+ylim(0,350)+
  geom_smooth(method='lm',color='red')+
  ggtitle("Free sulfur dioxide v/s Total sulfur dioxide")+
  ylab("Total sulfur dioxide")+
  xlab("Free sulfur dioxide")
```

This plot shows that with increase in free sulfur dioxide the proportion of
total sulfur dioxide increases.


```{r bivariate_alcohol_quality}

by(white_wine_df$alcohol,white_wine_df$rating,summary)

ggplot(aes(x=rating,y=alcohol),data=white_wine_df)+
   geom_jitter( alpha = .3)  +
  geom_boxplot( alpha = .5,color = 'blue')+
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4)+
  ggtitle("Alcohol v/s Quality")
```


This plot shows that good quality of wine had high median of 12.0 alcohol. This
visualisation shows that as alcohol content increase the quality of wine 
increases.

```{r bivarite_alcohol_quality_stats}

ggplot(aes(x=quality,y=alcohol),data=white_wine_df)+
  geom_point(alpha=1/10,position=position_jitter(h=0),color='red')+
  geom_line(stat='summary',fun.y=mean)+
  geom_line(stat='summary',fun.y=quantile, fun.args = list(probs = .1),
            linetype=2,color='blue')+
  geom_line(stat='summary',fun.y=quantile, fun.args = list(probs = .9),
            linetype=2,color='blue')+
  geom_line(stat='summary',fun.y=quantile, fun.args = list(probs = .5),
            color='blue')+
  ggtitle("Alcohol v/s quality summary")

```

This plot the summary statistic between alcohol and quality with black line
representing the mean, and the 0.9,0.5,0.1 quantile relationship between 
the two.



```{r bivariate_total_sulfur_density}

ggplot(aes(x=total.sulfur.dioxide,y=density),data=white_wine_df)+
  geom_point(alpha=1/10,position=position_jitter(h=0))+
  geom_smooth(method='lm',color='red')+
  ggtitle("Total sulfur dioxide v/s density")+
  xlab("Total sulfur dioxide")+
  ylab("Density")

```

This plot shows that with increase in total sulfur dioxide the density of white
wine increases.

```{r bivariate_chloride_density}

ggplot(aes(x=chlorides,y=density),data=white_wine_df)+
  geom_point(alpha=1/10,position=position_jitter(h=0))+
  xlim(0,0.15)+ylim(0.985,1.005)+
  geom_smooth(method='lm',color='red')+
  ggtitle("Chlorides v/s Density")

```

This plot shows that with increase in chlorides the density of white
alcohol as well increases.


```{r density_quality}

by(white_wine_df$chlorides,white_wine_df$rating,summary)

ggplot(aes(x=rating,y=chlorides),data=white_wine_df)+
  geom_boxplot()+
  ggtitle("Quality v/s Chlorides")+
  ylim(c(0,0.10))


```

This mean that with good quality of wine the proportion of chlorides is less
and for average quality of wine the proportion of chlorides in more. The 
chloride proportion increases with better quality of wine and then decreases.

```{r}

fit <- aov(chlorides ~ rating, data=white_wine_df)
#plot(fit)
summary(fit)

```

```{r alcohol_residual_sugar}

ggplot(aes(x=alcohol,y=residual.sugar),data=white_wine_df)+
  geom_point(alpha=1/10,position=position_jitter(h=0.1))+
  ylim(0,30)+
  geom_smooth(method='lm',color='red')+
  ggtitle("Alcohol v/s Residual sugar")+
  ylab("Residual sugar")

```

This plot shows that large number of wines have residual sugar content decrease
as alcohol proportion increases.


```{r alcohol_density}

ggplot(aes(x=alcohol,y=density),data=white_wine_df)+
  geom_point(alpha=1/10,position=position_jitter(h=0))+
  geom_smooth(method='lm',color='red')+
  ylim(0.985,1.005)+
  ggtitle("Alcohol v/s density")

```

This plot shows that large number of wines have density reduced as the alcohol
proportion increases



```{r bivariate_fixed_acidity_pH}

ggplot(aes(x=alcohol,y=total.sulfur.dioxide),data=white_wine_df)+
  geom_point(alpha=1/10,position=position_jitter(h=0.1))+
  geom_smooth(method='lm',color='red')+
  ggtitle("Alcohol v/s Total Sulfur dioxide")+
  ylab("Total Sulfur dioxide")

```

These plot show that as alcohol in white wine increase the total sulfur dioxide
content decreases.


# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

Here in this investigation I observed strong relation between alcohol and 
quality,residual sugar and density,free and total sulfur dioxide,fixed acidity 
and citric acid, chlorides and density.

The correlation constant values of other variables in the dataset wrt to 
quality are :

* citric.acid:quality = -0.009209091
* free.sulfur.dioxide:quality = 0.008158067
* fixed.acidity: quality= -0.1136628
* volatile.acidity:quality = -0.194723
* residual.sugar:quality = -0.09757683
* chlorides:quality = -0.2099344
* total.sulfur.dioxide:quality = -0.1747372
* density:quality = -0.3071233 
* pH:quality = 0.09942725
* sulphates:quality=0.05367788
* alcohol: quality=0.4355747

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

I observed positive correlation between free and total sulfur dioxide(0.615501),
alcohol and quality(0.4355747),Fixed acidity and Citric acid(0.2891807),
Chlorides and Density(0.2572113) and negative correlation between density
and quality(-0.3071233), Alcohol and Residual sugar(-0.4506312), 
Alcohol and Density(-0.7801376) and Fixed acidity and pH(-0.4258583)

### What was the strongest relationship you found?

I found the strongest positive correlation between residual sugar and density
with correlation constant value of 0.8389665.

# Multivariate Plots Section

Here in this section I will like to explore the relationship of 2 or more 
variables with respect to wine rating

In wine tasting, humans are least sensitive to the taste of sweetness 
(in contrast to sensitivity to bitterness or sourness) with the majority of the
population being able to detect sugar or "sweetness" in wines 
between 1% and 2.5% residual sugar. Additionally, other components of wine such 
as acidity and tannins can mask the perception of sugar in the wine.

First I will like to explore the relationship of residual sugar with density 
and does it impact the wine rating

```{r Multivariate_sugar_density_quality}

ggplot(aes(x = wine.type, y = residual.sugar/density), data = white_wine_df) + 
   geom_boxplot(aes(fill = rating))+
   ylab("Residual sugar per density")+
   xlab("Sugar content")+
  ggtitle("Relationship between Sugar content, 
           Residual sugar/density, Rating ")

```

This plot explores the relationship between wine sweetness to residual sugar
per density and how it impacts the wine rating. It shows that as sweetness level
of wines increase, the density as well increases. And it shows that sweet wines
have higher sugar/density and are just averagely rated.

Lets go upon exploring the relationship between free and total sulfur dioxide
with respect to the wine rating

```{r Multivariate_free_total_sulfur_quality}

ggplot(aes(x=free.sulfur.dioxide,y=total.sulfur.dioxide,color=rating),
       data=white_wine_df)+
  geom_point(alpha=1/5,position=position_jitter(h=0),stat = 'summary',fun.y=mean)+
    geom_smooth(method = "lm", se = FALSE,size=1)  +
   scale_color_brewer(palette = "Greens",type='seq',
                     guide=guide_legend(title='Rating'))+
  ggtitle("Free vs Total sulfur dioxide wrt to rating")+
  xlab("Free sulfur dioxide")+
  ylab("Total sulfur dioxide")+
  scale_fill_brewer(direction = -1) + theme_dark()
 

```

This plot shows that "Good" rated wines have lower proportion of free v/s total
sulfur dioxide proportion as compared to "Bad" wines.This plot shows that for
"Good" rated wines, the proportion of free sulfur dioxide vs total sulfur 
dioxide has to be less. As the wine quality increases the proportion of free vs 
total sulfur dioxide decreases

Acids are one of 4 fundamental traits in wine (the others are tannin, alcohol 
and sweetness). Acidity gives wine its tart and sour taste. SO let us go on
by exploring the total acidity(fixed+volatile) with respect to wine rating.

```{r Multivariate_total_acid_quality}

qplot(x=total.acidity,data=white_wine_df,
      geom='freqpoly', color=rating)+
  ggtitle("Total acidity v/s Rating")+xlab("Total acidity")

```

This plot shows the density curve for differently rated wines - Good,
Bad and average with respect to their total acidity. Here the density of total 
acidity of average rated wines is the maximum, followed by bad wines and then 
by good wines.  

**Citric acid** is a weak organic acid, which is often used as a natural 
preservative or additive to food or drink to add a sour taste to food.In terms 
of wines it is often added to wines to increase acidity,or to give a "fresh" 
flavor. Let us explore the relationship of citric acid with total and volatile
acidity and corresponding sugar level in the wines

```{r Multivariate_Plot_acidity_sugar}

ggplot(filter(white_wine_df,!is.na(citric_acid_bucket)),
       aes(total.acidity,volatile.acidity,color=citric_acid_bucket))+
   geom_point(alpha=1/5,position=position_jitter(h=0)) +
      geom_smooth(method = "lm", se = FALSE,size=1)  +
  facet_wrap(~ wine.type)+
  scale_colour_brewer(palette = 'Reds' ,type='seq',
                   guide=guide_legend(title='Citric acid'))+
  ylab("Volatile Acidity")+
  ggtitle("Relationship between total and volatile acidity, citric acid and 
        sugar content")+
  scale_fill_brewer(direction = -1) + theme_dark()

```

This multi variate scatter plot groups wines based on sugar content and citric 
acid content. For dry wines, most of citric acid content falls within bucket of 
(0,0.3] with low level of total acidity and medium volatile acidity. Dry wines 
also contain prominent wines with citric acid in bucket of (0.3,0.6] which are
scattered for volatile acidity and fixed acidity. Medium wines have higher
proportion of wines in (0.6,1.7] citric acid bucket as compared to dry wines. 
And for wines with content between 0-0.6 citric acid content is medium with 
total and volatile acidity proportion. 

Let us explore the relationship of citric acid with alcohol and total acidity 
and sweetness.

```{r Multivariate_alcohol_acid_citric_sugar}

ggplot(filter(white_wine_df,!is.na(citric_acid_bucket)),
       aes(alcohol,total.acidity, color=citric_acid_bucket))+
    geom_point(alpha=1/8,position=position_jitter(h=0)) +
      geom_smooth(method = "lm", se = FALSE,size=1)  +
  facet_wrap(~ wine.type)+
  scale_colour_brewer(palette = "Oranges",type='seq',
                   guide=guide_legend(title='Citric acid bucket'))+ 
  xlab("Alcohol content")+
  ylab("Total Acidity")+
  ggtitle("Relationship between alcohol, total acidity, 
          citric acid and sugar level ")+
  scale_fill_brewer(direction = -1) + theme_dark()

```

This plot shows  relationship among the variables i.e. alcohol content,
total acidity wrt to citric acid and sweetness in wine. This shows that dry 
wines have normally moderately low level of alcohol and low/medium level of  
citric acid. Medium sweeted wines also have moderately low level of alchol with
large number of wines containing medium level of citric acid.

Chloride concentrations in excess tend to give a full malty taste while Sulfate
ions, in contrast, tend to accentuate hop flavors and bitterness. So let us
analyze this relationship wrt to alcohol, the total acidity and sugar level to 
determine the balance in flavours

```{r Multivariate_sulfate_acid_alcohol_sugar_rating}

ggplot(aes(chloride_sulfate_ratio,total.acidity), data = white_wine_df) + 
   geom_boxplot(aes(fill = alcohol_content) )+
  facet_wrap(~ wine.type)+
  ylim(c(5,10))+
  xlab("Chloride sulfate ratio")+
 ylab("Total acidiy")+
   ggtitle("Relationship between chloride sulfate ratio, total acidity, alcohol 
           and wine type")

```

This plot explores a weak relationship between chloride sulfate ratio v/s total
acidity wrt to alcohol content and sweetness. Here dry wines are normally very
bitter with low level of alcohol and medium level of acidity. Medium sweeted
wines are also very bitter with low level of alcohol with high level of acidity.
Even sweet wines have high proprortion of chloride/sulfate ratio to balance out
the sweetness


```{r Multivariate_acidity_ph_sugar_rating}

ggplot(white_wine_df,aes(pH,total.acidity,color=wine.type))+
  geom_point(alpha=1/20,position=position_jitter(h=0)) +
  geom_smooth(method = "lm", se = FALSE,size=1)  +
  facet_wrap(~ rating)+
  scale_color_brewer(palette = 'Reds',type='seq',
                   guide=guide_legend(title='Sugar level'))+ 
  ylab("Total Acidity") +
  ggtitle("Relationship between pH, total acidity, wine type 
          and wine rating")+
  scale_fill_brewer(direction = -1) + theme_dark()

```

This plot explores the relationship of pH wrt to fixed acidity on basis of 
sweetness and wine rating. This shows that good wines haves low level of total
acidity and are normally dry/medium level of sweetness.And bad wines more
dry/medium level of sweetness as compared to good wines.

Let us determine the relationship if any present between total acidity, citric
acid, sweetness and wine rating

```{r  Multivariate_acid_citric_sugar_rating}

ggplot(white_wine_df,aes(total.acidity,citric.acid,color=wine.type))+
  geom_point(alpha=1/10,position=position_jitter(h=0)) +
  geom_smooth(method = "lm", se = FALSE,size=1)  +
  facet_wrap(~ rating)+
    scale_color_brewer(palette = 'Greens',type='seq',
                   guide=guide_legend(title='Wine rating'))+
  xlab("Total acidity")+
  ylab("Citric acid") +
  ggtitle("Relationship between Total acidity, citric acid, wine type and 
          wine rating")+
  scale_fill_brewer(direction = -1) + theme_dark()

```

This plot explores the relationship between citric and total acidity wrt to
sweetness and wine rating. Here for good wines, a very small proportion is
medium sweet and else are normally dry. The citric acid proportion also is less.
For bad wines, high percentage of wines are dry followed by medium wines. The 
citric acid level is high as compared to good wines.

```{r  Multivariate_acidity_alcohol_sugar_rating}

ggplot(white_wine_df,aes(total.acidity,alcohol,color=wine.type))+
 geom_point(alpha=1/10,position=position_jitter(h=0)) +
      geom_smooth(method = "lm", se = FALSE,size=1)  +
  facet_wrap(~ rating)+
  scale_color_brewer(palette='Oranges',type='seq',
                   guide=guide_legend(title='Wine rating'))+
  xlab("Total acidity")+
  ggtitle("Relationship between Total acidity, alcohol, wine type 
          and wine rating")+
  scale_fill_brewer(direction = -1) + theme_dark()

```

This plot explores the relationship between total acidity and alcohol wrt to 
wine sweetness and rating. A very low proportion falls into "good" wine 
category followed by bad rating. Good and bad wines have medium/dry level of 
sweetness with  alcohol in the range 8-14.

Today, the use of sulfur dioxide (SO2) is widely accepted as a useful 
winemaking aide. It is used as a preservative because of its anti-oxidative and 
anti-microbial properties in wine. Let us explore this relationship with wrt to
ph, sugar and wine rating.

```{r Multivariate_ph_sulfur_dioxide_sweet_ratig}

ggplot(white_wine_df,aes(pH,total.sulfur.dioxide,color=wine.type))+
 geom_point(alpha=1/10,position=position_jitter(h=0)) +
      geom_smooth(method = "lm", se = FALSE,size=1)  +
  scale_color_brewer(palette='Reds',type='seq',
                   guide=guide_legend(title='Wine rating'))+
  facet_wrap(~ rating)+
  ylab("Total sulfur dioxide")+
  ggtitle("Relationship between pH, total sulfur dioxide, wine type
          and wine rating")+
  scale_fill_brewer(direction = -1) + theme_dark()

```

This plot explores the relationship between pH and total sulfur dioxide wrt to
sweetness and wine rating. Here the good wines have average count of total 
sulfur dioixe with dry/medium level of sweetness. The bad wines have large 
number of dry wines with scattered pH and high total sulfur dioxide.

Let us explore a relationship between sugar level/ chloride sulphate ratio, 
alcohol level and how it impacts the wine rating

```{r Multivariate_chlorine_sulphate_sugar_alcohol_rating}

ggplot(aes(wine.type,chlorides/sulphates), data = white_wine_df) + 
   geom_boxplot(aes(fill = alcohol_content) )+
  facet_wrap(~ rating)+
    xlab("Sugar level")+
  ylim(c(0,0.2))+
  ylab("Chloride/sulphate ratio")+
   ggtitle("Relationship between sugar,chloride sulphate ratio,alcohol and
           rating")

```

From this plot we can see that good rated wines have a balance of sugar and 
saltiness with proportionate level of sugar and chloride/sulphate ratio and
have normally lower level level of alcohols. Bad wines have high level of sugar
and chloride/sulphates and low level of alcohols


# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

In this part of investigation, I explored relationship among multiple variables
int the data set. I found that as sugar content increases in wine,
the density increases linearly and the wine quality decreases.Also 
high  rated wines have lower proportion of free v/s total sulfur dioxide 
proportion as compared to poor rated wines.As total acid density increases in
wine the quality decreases, average rated wines are more dry as compared to bad
wines which are more mediumly sweeted.Good rated wines have more balance of
total acidity with citric acid ratio, dry wines have normally moderately 
low level of alcohol with high total acidity and medium level of citric acid.
Dry wines have normally moderately low level of alcohol and low/medium level of  
citric acid. Medium sweeted wines also have moderately low level of alchol with
large number of wines containing medium level of citric acid. From my analysis
I cannot pin point a single variable which could directly impact the quality,
as they say its more about balance! It has to be a balance of acidity, sweetness,
alcohol and sugar.

### Were there any interesting or surprising interactions between features?

An interesting observation made is that no single variable can by itself be 
responsible for the quality of the wine. Wines need to be a balanced in terms of
acidity, alcohol, sugar, saltiness etc. I found that a combination of medium 
total acidity and  dry/medium sweetness, low level of  citric acid, less of 
total sulfur dioxide, low/moderately low level of alcohol wrt to low level of
residual sugar and chloride/sulphate ratio would be atleast required to increase
the chance of that wine sample to be among the higher quality wine set.



To conclude I would like to add some inferential statistical tests to determine 
statistical significance of the various variables wrt  wine quality. **Analysis  
of variance (ANOVA)** is a collection of statistical models used to analyze the 
differences among group means and their associated procedures
(such as "variation" among and between groups). It is a quick, easy way to rule 
out un-needed variables that contribute little to the explanation of a dependent
variable. So let us determine the relationship of quality with newly added
categorical variables like alcohol_content, citric_acid_bucket,wine.type,
total_sulfur_dioxide_bucket and total.acidity

```{r annova_test}

wines <- aov(quality ~ alcohol_content*citric_acid_bucket*wine.type*
               total_sulfur_dioxide_bucket*total.acidity,data=white_wine_df)
summary(wines)

```

In most cases we put significance at the alpha=.05 level, or we require the 
P value to be less then .05 to be considered statistically significant. Based
upon the above results I would say to determine wine quality the following
relationships are significant wrt to quality :

* alcohol_content<br/>
* citric_acid_bucket                        
* wine.type                                  
* total.acidity                              
* citric_acid_bucket:wine.type              
* wine.type:total_sulfur_dioxide_bucket      
* citric_acid_bucket:total.acidity           
* wine.type:total.acidity                    
* total_sulfur_dioxide_bucket:total.acidity  



------

# Final Plots and Summary

### Plot One


```{r Plot_One}

by(white_wine_df$alcohol,white_wine_df$rating,summary)

ggplot(aes(x=rating,y=alcohol),data=white_wine_df)+
   geom_jitter( alpha = .3)  +
  geom_boxplot( alpha = .5,color = 'blue')+
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4)+
  ggtitle("Alcohol v/s Quality")


```


### Description One

I have chosen this plot as per correlation matrix for bivariate analysis, 
alcohol and quality had the highest correlation. This plot shows that "Good"" 
quality of wine had high median of 12.0 followed by "average"" for 11.8 and 
"bad" for 10.4 % level of aldohol. This visualisation shows that as alcohol 
content increase the quality of wine increases.


### Plot Two

```{r echo=FALSE,message=FALSE,warning=FALSE,Plot_Two}

ggplot(white_wine_df,aes(total.acidity,alcohol,color=wine.type))+
 geom_point(alpha = 0.3, h=0) +
      geom_smooth(method = "lm", se = FALSE,size=1)  +
  facet_wrap(~ rating)+
  scale_color_brewer(palette = 'RdBu',type='seq',
                   guide=guide_legend(title='Wine rating'))+
  xlab("Total acidity g / dm^3")+
  ylab("Alcohol %")+
  ggtitle("Relationship between Total acidity, alcohol, wine type 
          and wine rating")+
  scale_fill_brewer(direction = -1) + theme_dark()

```


### Description Two

This plot explores the relationship among various variables i.e. 
between total acidity and alcohol wrt to  wine sweetness and rating. Since 
alcohol,sweetness, acidity are some of the major characteristics which 
influence wine quality I wanted to explore this relationship. A very low 
proportion of the wines in the dataset fell into "good" wine category followed 
by "bad" rate in thedataset. Maximum  number of wines are averagely rated. It 
appears high level of alcohol is an important ingredient in wine preparation. 
Good wines have almost equal number of medium/dry sweeted wines with high level
of alcohol. Bad wines have large number of wine samples with medium/dry level 
of sweetness with  alcohol in the range 8-14.


### Plot Three
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_Three}

ggplot(aes(x = wine.type, y = residual.sugar/density), data = white_wine_df) + 
   geom_boxplot(aes(fill = rating))+
   ylab("Residual sugar per density")+
   xlab("Sugar content")+
  ggtitle("Relationship between Sugar content, 
           Residual sugar/density, Rating ")

```

### Description Three


This plot explores the relationship between residual sugar wrt to residual sugar
per density and how it impacts the wine rating. It shows that as sweetness level
of wines increase, the density as well increases. And it shows that sweet wines
have higher sugar/density and are just averagely rated while the dry wines
have low sugar/density and are either/average/good rated.

# Reflection

In this project I explored the white wine dataset which contained information 
about 4898 wines for 13 variables i.e X,fixed.acidity,volatile.acidity,
citric.acid, residual.sugar,chlorides,free.sulfur.dioxide,total.sulfur.dioxide,
density,pH,sulphates,alcohol,quality. I did exploratory analysis on the data
set for single, two and multiple variables in the dataset using R 
plots/visualizations.

In the first section of univariate analysis, I plotted the distribution of each
of the data set to better understand the sample data of white wines. I created
some additional variables such as rating based upon the wine quality,total 
acidity which was the sum of fixed and variable acidity, wine type which
determined the sugar level of the wines, grouped total sulfur dioxide and 
citric acid proportion into various buckets and chloride sulphate ratio.

Certain variables like fixed acidity,chloride and alcohol variables had to be 
log transformed to make the plot more normally distributed. Other than that 
citric acid, sulfur dioxide and residual sugar plots x- axis had to be 
re-adjusted to have a better understanding of the visualization.

Next in the bivariate section, I explored the correlation among the various
variables where correlation among residual sugar and density appeared the 
highest with alcohol and density appeared to be the weakest.The variables
having the highest correlation with wine quality were alcohol and the lowest
being chlorides.

In the multivariate analysis, I plotted couple of visualizations which 
explored the relationship of wine quality with residual sugar, alcohol, total
acidity, sulfur dioxide, chloride sulfate ratio,pH. An interesting observation 
made was that no single variable can by itself be responsible for the 
quality of the wine. Wines need to be a balanced in terms of acidity, alcohol, 
sugar, saltiness etc. I found that a combination of medium total acidity and 
dry/medium sweetness, low level of citric acid, less of total sulfur dioxide, 
low/moderately low level of alcohol wrt to low level of residual sugar and 
chloride/sulphate ratio would be atleast required to increase the chance of 
that wine sample to be among the higher quality wine set. 

The road-block which I hit in this exercise was choosing the dataset. Initially 
I was keen on exploring the presidential election dataset and was half way
through as well, but I hit a road block in between in the bivariate section
where I could not establish correlation among the categorical variables in
the section. So I thought of switching to the wine dataset. This was also not
a smooth sailing for me as I could not establish a strong correlation among
the variables in the multi variate section which directly influenced the wine
quality. So I read couple of articles in the Internet to better understand
wine preperation and the significance of various chemical constituents and then
I was able to move forward in my exploration.

Coming back to the white wine data set, I feel additional research about the 
data set needs to be carried out considering other factors as well. Because 
there can be other non chemical factors like  location,topography,soil,climate,
temperature which might influence the wine preperation and hence impact the 
wine quality.Future scope includes creation of a linear predictive model for 
wine analysis.