Explore White Wine Quality 
========================================================

### By Ruchita Maheshwary

In this report, I analyzed the data set for around 5000 white wines.

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(data.table)
library(dplyr)
library(gridExtra)
theme_update(axis.title=element_text(size=12),axis.text=element_text(size=10),
             plot.title=element_text(size=14))

```


```{r echo=FALSE, Load_the_Data}
# Load the Data

white_wine_df <-read.csv('wineQualityWhites.csv')

```

This data set is used to model wine preferences by data mining from 
physicochemical properties. In the dataset, the inputs include objective tests 
(e.g. PH values) and the output is based on sensory data (median of at least 3 
evaluations made by wine experts). Each expert graded the wine quality between 0
(very bad) and 10 (very excellent). 

**Brief description about the dataset variables:**

* fixed acidity: most acids involved with wine or fixed or nonvolatile (do not 
  evaporate readily)
* volatile acidity: the amount of acetic acid in wine, which at too high of 
  levels can lead to an unpleasant, vinegar taste
* citric acid: found in small quantities, citric acid can add 'freshness' and 
  flavor to wines
* residual sugar: the amount of sugar remaining after fermentation stops, it's 
  rare to find wines with less than 1 gram/liter and wines with greater than 45 
  grams/liter are considered sweet
* chlorides: the amount of salt in the wine
* free sulfur dioxide: the free form of SO2 exists in equilibrium between 
  molecular SO2 (as a dissolved gas) and bisulfite ion; it prevents microbial 
  growth and the oxidation of wine
* total sulfur dioxide: amount of free and bound forms of S02; in low 
  concentrations, SO2 is mostly undetectable in wine, but at free SO2 
  concentrations over 50 ppm, SO2 becomes evident in the nose and taste of wine
* density: the density of water is close to that of water depending on the 
  percent alcohol and sugar content
* pH: describes how acidic or basic a wine is on a scale from 0 (very acidic) 
  to 14 (very basic); most wines are between 3-4 on the pH scale
* sulphates: a wine additive which can contribute to sulfur dioxide gas (S02) 
  levels, wich acts as an antimicrobial and antioxidant
* alcohol: the percent alcohol content of the wine
* quality (score between 0 and 10): Output variable (based on sensory data)

# Univariate Plots Section

Below is a brief summary of the data:

```{r echo=FALSE, Univariate_Plots}
dim(white_wine_df)

str(white_wine_df)

summary(white_wine_df)

```


```{r echo=FALSE, warning=FALSE,message=FALSE, Univariate_Plots_Fixed_Acidity}

summary(white_wine_df$fixed.acidity)

# Plot for fixed acidity

  qplot(x=fixed.acidity,data=white_wine_df,binwidth=0.1,xlab = "Fixed Acidity")+
    scale_x_continuous(limits=c(3,15),breaks=seq(3,15,3))+
    ggtitle("Fixed Acidity per g / dm^3")
  
```

White wine's fixed acidity plot is a normal distribution with highest 
concentration of fixed acidity at 6.8. Around 50% of wines have fixed acidity 
more than 6.8 and around top 25% lie in the range 6.8-7.3 with an outlier at
14.2.


```{r echo=FALSE, warning=FALSE,message=FALSE, Univariate_Plots_Volatile_Acidity}

summary(white_wine_df$volatile.acidity)

qplot(x=volatile.acidity,data=white_wine_df,binwidth=0.01,xlab="Volatile Acidity")+
  scale_x_continuous(limits=c(0,1.2),breaks = seq(0,1.2,0.5))+
    ggtitle("Volatile Acidity per g / dm^3")


```

The summary statistics show that maximum volatile acidity is at 0.26. 
Around 50% of wines have volatile acidity more than 0.26 and around 
top 25% lie in the range 0.26-0.32 with an outlier at
1.1.

A log transformation is required to transform the log tail to make this plot 
more normally distributed.


```{r echo=FALSE, warning=FALSE, message=FALSE,Univariate_Plots_Volatile_Acidity_Log_Transform}

ggplot(aes(x=volatile.acidity),data=white_wine_df)+
  geom_histogram()+
  scale_x_log10()+
  xlab("Volatile Acidity")
    ggtitle("Volatile Acidity Histogram with log transformation")

```



```{r echo=FALSE,warning=FALSE, message=FALSE, Univariate_plot_citric_acid_transform}

summary(white_wine_df$citric.acid)

qplot(x=citric.acid,data=white_wine_df,binwidth=0.02,xlab = "Citric Acid")+
    scale_x_continuous(limits=c(0,2),breaks=seq(0,2.0,0.5))+
  ggtitle("Citric acid concentration per g / dm^3")


```

The plot has long tailed which after re - adjusting the axis could be normally
distributed. Here the median of citric acid concentration is 0.32 g/dm^3

```{r echo=FALSE,warning=FALSE, message=FALSE, Univariate_plot_citric_acid}

qplot(x=citric.acid,data=white_wine_df,binwidth=0.01,xlab = "Citric Acid")+
    scale_x_continuous(limits=c(0,1),breaks=seq(0,1,0.2))+
  ggtitle("Citric acid concentration per g / dm^3")

```


```{r echo=FALSE, message=FALSE, warning=FALSE,Univariate_plot_residual_sugar}

summary(white_wine_df$residual.sugar)

 qplot(x=residual.sugar,data=white_wine_df,binwidth=0.1,xlab = "Residual Sugar")+
     scale_x_continuous(limits=c(0,70),breaks=seq(0,70,10))+
   ggtitle("Residual Sugar per g / dm^3")

```

Here the median residual sugar is 5.2 g/dm^3 with an outlier at 65.8.
We can perform readjust the plot to get a better understanding of it.


```{r echo=FALSE, message=FALSE, warning=FALSE,Univariate_plot_residual_sugar_log}

 qplot(x=residual.sugar,data=white_wine_df,binwidth=0.1,xlab = "Residual Sugar")+
     scale_x_continuous(limits=c(0,20),breaks = seq(0,20,5))+
   ggtitle("Residual Sugar Re-adjusted Plot")

```

This visualisation  of residual sugar appears skewed to the right with a 
high peak at around 3.


```{r echo=FALSE, warning=FALSE, message=FALSE, Univariate_plot_chlorides}

summary(white_wine_df$chlorides)

qplot(x=chlorides,data=white_wine_df,binwidth=0.001)+
  scale_x_continuous(limits = c(0,0.35),breaks = seq(0,0.35,0.1))+
  ggtitle("Chloride content per g / dm^3")
```

This visualization has a long tail with an outlier at around 0.34. We can
apply a log transformation to get a better understanding of this plot.

```{r echo=FALSE, warning=FALSE, message=FALSE, Univariate_plot_chlorides_log}

ggplot(aes(x=chlorides),data=white_wine_df,binwidth=0.1)+
  geom_histogram()+
  scale_x_log10()+
  scale_x_continuous(limits=c(0,0.15),breaks=seq(0,0.15,0.3))+
  ggtitle("Chloride content after log transformation")
```

After log transformation this visualization still appears skewed towards the right.

```{r echo=FALSE,warning=FALSE, message=FALSE,Univariate_plot_free_sulfur_dioxide}

summary(white_wine_df$free.sulfur.dioxide)

qplot(x=free.sulfur.dioxide,data=white_wine_df,binwidth=2,xlab = "Free sulfur dioxide")+
  scale_x_continuous(limits = c(0,300),breaks = seq(0,300,50))+
  ggtitle("Free sulfur dioxide per g/dm^3")

```

This graph has a median of free sulfur dioxide of 34 g/dm^3 and have an outlier
at 289. On re-adjusting the axis to obtain an understanding of the shape of 
the plot.

```{r echo=FALSE,warning=FALSE, message=FALSE,Univariate_plot_free_sulfur_dioxide_readjust}

qplot(x=free.sulfur.dioxide,data=white_wine_df,binwidth=2,
      xlab = "Free sulfur dioxide ")+
  scale_x_continuous(limits = c(0,100),breaks = seq(0,100,10))+
  ggtitle("Free sulfur dioxide per g/dm^3")

```

After readjusting the axis to remove the outlier and after limit the range of 
x-axis to 0-100 g/dm^3, the plot is clearly a normal distribution. 

```{r echo=FALSE, warning=FALSE,message=FALSE,Univariate_plot_total_sulfur_dioxide}

summary(white_wine_df$total.sulfur.dioxide)

qplot(x=total.sulfur.dioxide,data=white_wine_df,binwidth=5,
      xlab="Total sulfur dioxide")+
  scale_x_continuous(limits = c(0,450),breaks = seq(0,450,50))+
  ggtitle("Total sulfur dioxide per g/dm^3")


```

This plot is normally distributed with an outlier at 440 g/dm^3 with maximum
concentration of sulfur dioxide at around 134 g/dm^3.


```{r echo=FALSE, warning=FALSE,message=FALSE,Univariate_plot_density}

summary(white_wine_df$density)

qplot(x=density,data=white_wine_df,binwidth=0.0005)+
  scale_x_continuous(limits = c(0.98,1.01),breaks = seq(0.98,1.01,0.01))+
  ggtitle("Density per g/cm^3")

```

This plot had large number of outliers after 1.01 g/cm^3 which had been removed
after re-adjusting the x-axis. Now the plot appears to be normally distributed.
The median of density had been at 0.9937 g/cm^3.   

```{r echo=FALSE,warning=FALSE,message=FALSE, Univariate_plot_pH}

summary(white_wine_df$pH)

qplot(x=pH,data=white_wine_df,binwidth=0.02)+
  scale_x_continuous(limits = c(2.7,4),breaks = seq(2.7,4,0.5))+
  ggtitle("pH level")

```

This plot of pH level in the white wines is normally distributed with
the median at 3.188. Around 50% of white wines had pH level of more than 3.188
with the maximum being at 3.820.


```{r echo=FALSE,warning=FALSE, message=FALSE,Univariate_plot_sulphates}

summary(white_wine_df$sulphates)

qplot(x=sulphates,data=white_wine_df,binwidth=0.02)+
  scale_x_continuous(limits = c(0.2,1.1),breaks = seq(0,1.1,0.25))+
  ggtitle("Potassium sulphate per g/dm^3")

```

This plot is skewed to the right with a long tail and having an outlier at 1.08.
The median of sulphate concentration is 0.47 g/dm^3.

```{r echo=FALSE,warning=FALSE, message=FALSE,Univariate_plot_alcohol}

summary(white_wine_df$alcohol)

qplot(x=alcohol,data=white_wine_df,binwidth=0.1)+
  scale_x_continuous(limits = c(8,15),breaks = seq(8,15,1))+
  ggtitle("Alcohol content")

```

This plot appears to be skewed towards the right, after applying 
log transformations the plot appears to be bimodal.

```{r echo=FALSE,warning=FALSE, message=FALSE,Univariate_plot_alcohol_log}

qplot(x=alcohol,data=white_wine_df)+
  scale_x_log10()+
  ggtitle("Alcohol content")

```


```{r echo=FALSE, warning=FALSE,message=FALSE,Univariate_plot_quality}

summary(white_wine_df$quality)

ggplot(aes(x=quality),data=white_wine_df)+
  geom_bar()+
  scale_x_continuous(breaks=seq(3,9,1))

```

```{r echo=FALSE, warning=FALSE,message=FALSE,Univariate_wine_quality_calculations }

#Determining proportion of wine quality having rating 9 and 6.
wine_quality_9<-subset(white_wine_df,quality==9)
wine_quality_9_prop<-nrow(wine_quality_9)/nrow(white_wine_df)

wine_quality_6<-subset(white_wine_df,quality==6)
wine_quality_6_prop<-nrow(wine_quality_6)/nrow(white_wine_df)

#Adding extra variable for rating of the wine quality with 3-5 for bad,
# 6-7 for average and 8-9 as good

white_wine_df$rating <- NULL
white_wine_df$rating <-ifelse(white_wine_df$quality>=3 & white_wine_df$quality<=5,
                              'Bad',
                              ifelse(white_wine_df$quality>=6 & white_wine_df$quality<=7,
                              'Average','Good'))

white_wine_df$rating <- factor(white_wine_df$rating)

```


This bar plot shows that the median of wine quality has been 6 forming around
44.8% of the dataset with the best quality of wine having quality as 9 which 
forms around 1.02% of the dataset.



# Univariate Analysis

### What is the structure of your dataset?

There are 4898 observations of white wine in this dataset with 13 variables.
The first variable 'X' is a unique identifier for each record so it is ignored
in my analysis. Other variables like 	fixed.acidity, volatile.acidity,
citric.acid,residual.sugar,chlorides,free.sulfur.dioxide,total.sulfur.dioxide,
density,pH,sulphates,alcohol are all numeric variables with	quality 
being an integer variable.

### What is/are the main feature(s) of interest in your dataset?

My main point of interest is determing the factors which determine the quality
of white wine.

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

I would need to compute the correlation of each of the variables like 
fixed.acidity, volatile.acidity,citric.acid,residual.sugar,chlorides,
free.sulfur.dioxide,total.sulfur.dioxide, density,pH,sulphates,alcohol with the 
quality to determine which variables influence the quality and hence the 
taste of white wine. I will explore these relationships in the later sections
of my report.


### Did you create any new variables from existing variables in the dataset?

I added one variable called as rating in the dataset which is calculated on
the basis of the quality of the white wine. With 3-5 for "Bad",  6-7 for 
"Average" and 8-9 as "Good" rating. This factored variable will help me in
later investigating the relationship of variables with the wine quality


### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?

The fixed acidity,chloride and alcohol variables had to be log transformed
to make the plot more normally distributed. Other than that citric acid, 
sulfur dioxide and residual sugar plots x- axis had to  be re adjusted 
to have a better understanding of the visualization.


# Bivariate Plots Section

```{r echo=FALSE,message=FALSE, warning=FALSE, include=FALSE,correlation_values}


cor.test(white_wine_df$quality, white_wine_df$citric.acid,method = 'pearson')
#-0.009209091
cor.test(white_wine_df$quality, white_wine_df$free.sulfur.dioxide,method = 'pearson')
#0.008158067
cor.test(white_wine_df$quality, white_wine_df$fixed.acidity,method = 'pearson')
#-0.1136628
cor.test(white_wine_df$quality, white_wine_df$volatile.acidity,method = 'pearson')
#-0.194723
cor.test(white_wine_df$quality, white_wine_df$residual.sugar,method = 'pearson')
#-0.09757683
cor.test(white_wine_df$quality, white_wine_df$chlorides,method = 'pearson')
#-0.2099344
cor.test(white_wine_df$quality, white_wine_df$total.sulfur.dioxide,method = 'pearson')
#-0.1747372
cor.test(white_wine_df$quality, white_wine_df$density,method = 'pearson')
#-0.3071233 
cor.test(white_wine_df$quality, white_wine_df$pH,method = 'pearson')
#0.09942725
cor.test(white_wine_df$quality, white_wine_df$sulphates,method = 'pearson')
#0.05367788
cor.test(white_wine_df$quality, white_wine_df$alcohol,method = 'pearson')
#0.4355747

```

There are negative correlation between the following variables wrt quality

* fixed.acidity and quality(-0.1136628)
* volatile.acidity and quality(-0.194723)
* chlorides and quality(-0.2099344)
* total sulfur dioxide and quality(-0.1747372)
* density and quality(-0.3071233)


Negligible correlation: <br/>
* residual sugar and quality (-0.09757683)<br/>
* citric acid and quality(-0.009209091)<br/>
* free sulfur dioxide and quality(0.008158067)<br/>

Positive correlation :<br/>
* pH and quality(0.09942725)<br/>
* alcohol and quality(0.4355747)<br/>
* sulfates and quality(0.05367788)<br/>


Other correlations between these variables are:


```{r echo=FALSE,message=FALSE, warning=FALSE,include=FALSE,other_correlation}

cor.test(white_wine_df$fixed.acidity, white_wine_df$volatile.acidity,method = 'pearson')
#-0.02269729
cor.test(white_wine_df$alcohol, white_wine_df$citric.acid,method = 'pearson')
#-0.07572873
cor.test(white_wine_df$free.sulfur.dioxide, white_wine_df$total.sulfur.dioxide,method = 'pearson')
#0.615501
cor.test(white_wine_df$citric.acid, white_wine_df$pH,method = 'pearson')
#-0.1637482
cor.test(white_wine_df$citric.acid, white_wine_df$residual.sugar,method = 'pearson')
#0.09421162 
cor.test(white_wine_df$alcohol, white_wine_df$residual.sugar,method = 'pearson')
#-0.4506312
cor.test(white_wine_df$fixed.acidity, white_wine_df$citric.acid,method = 'pearson')
#0.2891807
cor.test(white_wine_df$chlorides, white_wine_df$sulphates,method = 'pearson')
#0.01676288
cor.test(white_wine_df$chlorides, white_wine_df$residual.sugar,method = 'pearson')
#0.08868454 
cor.test(white_wine_df$chlorides, white_wine_df$density,method = 'pearson')
#0.2572113
cor.test(white_wine_df$alcohol, white_wine_df$density,method = 'pearson')
#-0.7801376

cor.test(white_wine_df$pH, white_wine_df$density,method = 'pearson')
#-0.09359149

cor.test(white_wine_df$pH, white_wine_df$fixed.acidity,method = 'pearson')
#-0.4258583

cor.test(white_wine_df$residual.sugar, white_wine_df$density,method = 'pearson')
#0.8389665

```


From these analystics the top most positive and negative correlation 
coefficients, r based on strength of association[1] are 

* Residual sugar: density = 0.8389665
* Free sulfur dioxide: Total sulfur dioxide=0.615501
* Alcohol: quality = 0.4355747
* Fixed acidity:Citric acid=0.2891807
* Chlorides:Density=0.2572113
* Density:quality= -0.3071233
* Alcohol:Residual sugar=-0.4506312
* Alcohol:Density= -0.7801376
* Fixed.acidity:pH= -0.4258583

I will exploring these relationships in this section of the report.


```{r message=FALSE,warning=FALSE,echo=FALSE,bivariate_sugar_density}

ggplot(aes(x=density,y=residual.sugar),data=white_wine_df)+
  geom_point(alpha=1/10,position=position_jitter(h=0))+
  xlim(0.985,1.005)+ylim(0,30)+
   geom_smooth(method='lm',color='red')+
  ggtitle("Residual sugar v/s density")
```

This plot shows that as residual sugar in white wine increases, the density
increases.


```{r message=FALSE,warning=FALSE,echo=FALSE,bivariate_sulfur_dioxide}

ggplot(aes(x=free.sulfur.dioxide,y=total.sulfur.dioxide),data=white_wine_df)+
  geom_point(alpha=1/10,position=position_jitter(h=0.1))+
  xlim(0,100)+ylim(0,350)+
   geom_smooth(method='lm',color='red')+
  ggtitle("Free sulfur dioxide v/s Total sulfur dioxide")
```

This plot shows that with increase in free sulfur dioxide the proportion of
total sulfur dioxide increases.


```{r message=FALSE,warning=FALSE,echo=FALSE,bivariate_alcohol_quality}

by(white_wine_df$alcohol,white_wine_df$rating,summary)

ggplot(aes(x=rating,y=alcohol),data=white_wine_df)+
  geom_boxplot()+
 # geom_point(alpha=1/10,position=position_jitter(h=0))+
  #geom_smooth(method='lm',color='red')+
  ggtitle("Alcohol v/s Quality")
```


This plot shows that good quality of wine had high median of 12.6 alcohol. This
visualisation shows that as alcohol content increase the quality of wine 
increases.

```{r echo=FALSE,message=FALSE, warning=FALSE,bivarite_alcohol_quality_summary_statistics}

ggplot(aes(x=quality,y=alcohol),data=white_wine_df)+
  geom_point(alpha=1/10,position=position_jitter(h=0),color='red')+
  geom_line(stat='summary',fun.y=mean)+
   geom_line(stat='summary',fun.y=quantile, fun.args = list(probs = .1),
             linetype=2,color='blue')+
   geom_line(stat='summary',fun.y=quantile, fun.args = list(probs = .9),
             linetype=2,color='blue')+
   geom_line(stat='summary',fun.y=quantile, fun.args = list(probs = .5),color='blue')+
  ggtitle("Alcohol v/s quality summary")

```

This plot the summary statistic between alcohol and quality with black line
representing the mean, and the 0.9,0.5,0.1 quantile relationship between 
the two.



```{r message=FALSE,warning=FALSE,echo=FALSE,bivariate_fixed_citric_acidity}

ggplot(aes(x=fixed.acidity,y=citric.acid),data=white_wine_df)+
  geom_point(alpha=1/10,position=position_jitter(h=0))+xlim(3.5,10)+ylim(0,1)+
  geom_smooth(method='lm',color='red')+
  ggtitle("Fixed Acidity v/s Citric Acid")
```

This plot shows that with increase in fixed acidity the proportion of
citric acid increases.

```{r message=FALSE,warning=FALSE,echo=FALSE,bivariate_chloride_density}

ggplot(aes(x=chlorides,y=density),data=white_wine_df)+
  geom_point(alpha=1/10,position=position_jitter(h=0))+
  xlim(0,0.15)+ylim(0.985,1.005)+
  geom_smooth(method='lm',color='red')+
  ggtitle("Chlorides v/s Density")

```

This plot shows that with increase in chlorides the density of white
alcohol as well increases.


```{r message=FALSE,warning=FALSE,echo=FALSE,density_quality}

by(white_wine_df$chlorides,white_wine_df$rating,summary)

ggplot(aes(x=rating,y=chlorides),data=white_wine_df)+
  geom_boxplot()+
  #geom_point(alpha=1/10,position=position_jitter(h=0))+
  #ylim(0,0.15)+
  ggtitle("Quality v/s Chlorides")


```

This mean that with good quality of wine the proportion of chlorides is less
and for average quality of wine the proportion of chlorides in more. The 
chloride proportion increases with better quality of wine and then decreases.


```{r message=FALSE,warning=FALSE,echo=FALSE,alcohol_residual_sugar}

ggplot(aes(x=alcohol,y=residual.sugar),data=white_wine_df)+
  geom_point(alpha=1/10,position=position_jitter(h=0.1))+
  ylim(0,30)+
  geom_smooth(method='lm',color='red')+
  ggtitle("Alcohol v/s Residual sugar")

```

This plot shows that large number of wines have residual sugar content decrease
as alcohol proportion increases.


```{r message=FALSE,warning=FALSE,echo=FALSE,alcohol_density}

ggplot(aes(x=alcohol,y=density),data=white_wine_df)+
  geom_point(alpha=1/10,position=position_jitter(h=0))+
  geom_smooth(method='lm',color='red')+
  ylim(0.985,1.005)+
  ggtitle("Alcohol v/s density")

```

This plot shows that large number of wines have density reduced as the alcohol
proportion increases



```{r message=FALSE,warning=FALSE,echo=FALSE,bivariate_fixed_acidity_pH}

ggplot(aes(x=fixed.acidity,y=pH),data=white_wine_df)+
  geom_point(alpha=1/10,position=position_jitter(h=0.1))+
  geom_smooth(method='lm',color='red')+
  ggtitle("Fixed Acidity v/s pH")

```

These plot show that as fixed value in white wine increase the 
pH value decreases.


# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

Here in this investigation I observed strong relation between alcohol and quality,
residual sugar and density,free and total sulfur dioxide,fixed acidity 
and citric acid, chlorides and density.

The correlation constant values of other variables in the dataset wrt to 
quality are :

* citric.acid:quality = -0.009209091
* free.sulfur.dioxide:quality = 0.008158067
* fixed.acidity: quality= -0.1136628
* volatile.acidity:quality = -0.194723
* residual.sugar:quality = -0.09757683
* chlorides:quality = -0.2099344
* total.sulfur.dioxide:quality = -0.1747372
* density:quality = -0.3071233 
* pH:quality = 0.09942725
* sulphates:quality=0.05367788

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

I observed positive correlation between free and total sulfur dioxide(0.615501),
alcohol and quality(0.4355747),Fixed acidity and Citric acid(0.2891807),
Chlorides and Density(0.2572113) and negative correlation between density
and quality(-0.3071233), Alcohol and Residual sugar(-0.4506312), 
Alcohol and Density(-0.7801376) and Fixed acidity and pH(-0.4258583)

### What was the strongest relationship you found?

I found the strongest positive correlation between residual sugar and density
with correlation constant value of 0.8389665.

# Multivariate Plots Section

> **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

> **Tip**: Don't forget to remove this, and the other **Tip** sections before
saving your final work and knitting the final report!